---
title: "6101 Project 1"
author: Jeffrey Hu, Aaron Rock, Sanchit Vijay
date: 9/27/22 to 10/27/22
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r init, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
# Make sure that install.packages("matrixStats"), install.packages("reshape2"), install.packages("reshape") are installed

library(ezids)
library(ggplot2)
library(tidyverse)
library(MASS)
library(reshape)
library(dplyr)
library(matrixStats)
library(zoo)
library(ggfortify)
library(reshape2)
library(hexbin)
# knitr::opts_chunk$set(warning = F, results = "markup", message = F)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```

## Summary Paper Instructions
Write a roughly 10-page (definitely no more than 4000 words, charts do not count)
summary of the research and EDA process of your project. The summary should be
prepared in Rmarkdown and knitted into HTML. You may take some of the work in part
II (such as graphs and results) to include here. They can overlap. You may also make
modifications to your analysis based on feedback from your presentation. This summary is
to be presented to your boss, your client, or to be submitted for publication in journals.
Potential area of topics to address in this summary may include:
 What do we know about this dataset?
 What are the limitations of the dataset?
 How was the information gathered?
 What analysis has already been completed related to the content in your dataset?
 How did the research you gathered contribute to your question development?
 What additional information would be beneficial?
 How did your question change, if at all, after Exploratory Data Analysis?
 Based on EDA can you begin to sketch out an answer to your question?
 References (APA style preferred)

## Load the file
```{r}
zillow_prices = data.frame(read.csv("price.csv"))
zillow_focus = zillow_prices[c(1:5,7,13,17,23,24,26,113,162),]
zillow_focus = zillow_focus[c(-1,-3,-4,-5,-6)]
zillow_focus
```


## Clean and Transpose
Our cities of interest are NYC (Queens) (only data from December 2011 Onwards), LA, PA, Houston (Harris), Chicago (Cook), Dallas, Las Vegas, San Fran, Detroit (Wayne), Roanoke, Richmond, District of Columbia, Baltimore, Seattle(King). Here, we clean data set and transpose the dates/cities
```{r, results="markup"}
rownames(zillow_focus) = zillow_focus$City
zillow_focus = zillow_focus[c(-1)]
zillow_focus = as.data.frame(t(zillow_focus))
names(zillow_focus)<-str_replace_all(names(zillow_focus), c(" " = "" , "," = ""))
head(zillow_focus)
str(zillow_focus)
```

Turn date into a column
```{r, results="markup"}
zillow_focus <- cbind(Date = rownames(zillow_focus), zillow_focus)
rownames(zillow_focus) <- 1:nrow(zillow_focus)
head(zillow_focus)
```


## Convert the Date column from chr to date, and Total
Here, we are converting strings to dates and to %d/%m/%Y form with 01 for %d
```{r, results="markup"}
zillow_focus_test = zillow_focus
zillow_focus_test$Date <- str_replace_all(zillow_focus_test$Date, "[.]", " ")
zillow_focus_test$Month <- str_extract(zillow_focus_test$Date, "(\\w+)")
zillow_focus_test$Year <- str_extract(zillow_focus_test$Date, "\\w+$")
tmp_date = paste(match(zillow_focus_test$Month, month.name), zillow_focus_test$Year, sep="/")
tmp_date = paste("01", tmp_date, sep="/")
tmp_date = as.Date(tmp_date, "%d/%m/%Y")
zillow_focus$Date <- tmp_date
head(zillow_focus)
str(zillow_focus)
```


Add a total column, not including New York (which has NAs)
```{r, results="markup"}
zillow_total = zillow_focus
zillow_total$Total <- rowSums(zillow_total[3:14])
head(zillow_total)
```


## Line plot of total
Line plots of time series, one with only time
```{r}
ggplot(zillow_total, aes(x=Date, y=Total)) + geom_point() + geom_line() + labs(x="Time", y="Total Rent Prices", title="Plot of City Rent Change over Time")
```


## Plots of the Cities and Prices
Melt the ggplot by the cities
```{r, results="markup"}
zillow_melt <- melt(zillow_focus, id = "Date")
names(zillow_melt)[2] <- "City"
names(zillow_melt)[3] <- "Prices"
head(zillow_melt)
```


I'm not sure if histograms would be meaningful in the context of my part, but here's one of all the cities melted together, ignoring New York again. We can do more if you guys believe it's necessary. There is a huge skew because the bigger/more expensive cities chosen would have higher rent prices due to a difference in the cost of living.
I'll also do the Mean/Median/Quartiles of the overall plots here.
```{r, results="markup"}
ggplot(zillow_melt, aes(x=Prices, na.rm=TRUE)) + geom_histogram(binwidth=300)
ggplot(zillow_melt, aes(x=Prices, na.rm=TRUE)) + geom_boxplot()
summary(zillow_melt$Prices)
```

Lines of the time series over their changes.
```{r}
ggplot(zillow_melt, aes(x=Date, y=Prices, group=City, color=City, na.rm=TRUE)) + geom_point() + geom_line() + labs(x="Time", y="Rent Price", title="Plot of City Rent Prices over Time")
```


## First Order change plots
Find the first order change, difference between each month
```{r, results="markup"}
zillow_change <- as.data.frame(lapply(zillow_focus[2:14], diff, lag=1))
zillow_change$Date <- zillow_focus$Date[2:75]
zillow_change <- zillow_change %>% dplyr::select(Date, everything())
head(zillow_change)
```


Melt, and line graph of the changes for each city.
The graph is kind of hard to visualize, but it looks like San Francisco had the highest change from time to time.
```{r}
zillow_cmelt <- melt(zillow_change, id = "Date")
names(zillow_cmelt)[2] <- "City"
names(zillow_cmelt)[3] <- "PriceChange"
ggplot(zillow_cmelt, aes(x=Date, y=PriceChange, group=City, color=City, na.rm=TRUE)) + geom_point() + geom_line() + labs(x="Time", y="Price Change", title="Plot of City Rent Change over Time")
```

Cumulative change, to make the trends more clear in respect to each other
```{r, results="markup"}
zillow_cumulative <- as.data.frame(cumsum(zillow_change[2:14]))
zillow_cumulative$Date <- zillow_focus$Date[2:75]
zillow_cumulative <- zillow_cumulative %>% dplyr::select(Date, everything())
head(zillow_cumulative)
```


Another graph, this time of cumulative change. San Francisco had the greatest change, while Seattle had the second largest. Detroit had the most negative change, falling below 0 overall.
```{r}
zillow_cumumelt <- melt(zillow_cumulative, id = "Date")
names(zillow_cumumelt)[2] <- "City"
names(zillow_cumumelt)[3] <- "CumulativePriceChange"
ggplot(zillow_cumumelt, aes(x=Date, y=CumulativePriceChange, group=City, color=City, na.rm=TRUE)) + geom_point() + geom_line() + labs(x="Time", y="Cumulative Change", title="Plot of Cumulative City Rent Change over Time")
```


## First Order change, as a percentage proportion of the total rent
Find the first order change as a Percent
```{r, results="markup"}
zillow_proportion <- zillow_change[2:14]/zillow_focus[c(2:75), 2:14]*100
zillow_proportion$Date <- zillow_focus$Date[2:75]
zillow_proportion <- zillow_proportion %>% dplyr::select(Date, everything())
head(zillow_proportion)
```

It looks like Richmond has a pretty varied proportional change. It's still hard to make anything out of it, so let's move to the cumulative proportional change again.
```{r}
zillow_pmelt <- melt(zillow_proportion, id = "Date")
names(zillow_pmelt)[2] <- "City"
names(zillow_pmelt)[3] <- "ProportionChange"
ggplot(zillow_pmelt, aes(x=Date, y=ProportionChange, group=City, color=City, na.rm=TRUE)) + geom_point() + geom_line() + labs(x="Time", y="Proportional Change", title="Plot of City Rent Change over Time")
```


Quick histogram of the distribution of proportional changes. The proportional changes over time look pretty normal taken together as a whole, with maybe a slight right skew, which makes sense as we would guess that rent prices should generally increase over time with inflation and what not. Remember, these changes are in percentages. There are outliers on both sides of the boxplot, which matches what we see in the histogram. Here are the measures of variance and qqnorm plots as well
```{r, results="markup"}
ggplot(zillow_pmelt, aes(x=ProportionChange, na.rm=TRUE)) + geom_histogram() + labs(x = "Proportional Change in Percentages")
ggplot(zillow_pmelt, aes(x=ProportionChange, na.rm=TRUE)) + geom_boxplot()
qqnorm(zillow_pmelt$ProportionChange)
summary(zillow_pmelt$ProportionChange)
var(zillow_pmelt$ProportionChange, na.rm=TRUE)
sd(zillow_pmelt$ProportionChange, na.rm=TRUE)
```

Just making the cumulative melt for proportions now. We will drop New York before plotting the values, as there are too many NAs.

```{r, results="markup"}
zillow_pdrop <- zillow_proportion[3:14]
zillow_pcumulative <- as.data.frame(cumsum(zillow_pdrop))
zillow_pcumulative$Date <- zillow_focus$Date[2:75]
zillow_pcumulative <- zillow_pcumulative %>% dplyr::select(Date, everything())
head(zillow_pcumulative)
```

We will drop New York before plotting the values, as there are too many NAs.
Now, it looks like the proportional change of Seattle is a bit higher than San Francisco, and Richmond is up there too. Detroit still has the most negative proportional change overall.

```{r}
zillow_pcumumelt <- melt(zillow_pcumulative, id = "Date")
names(zillow_pcumumelt)[2] <- "City"
names(zillow_pcumumelt)[3] <- "CumulativeProportionChange"
ggplot(zillow_pcumumelt, aes(x=Date, y=CumulativeProportionChange, group=City, color=City, na.rm=TRUE)) + geom_point() + geom_line() + labs(x="Time", y="Cumu. Prop. Change", title="Plot of Cumulative City Rent Change over Time")
```

Box plot of each city's proportional change. As expected, San Francisco and Roanoke seem to have the largest ranges and variances. We are also dropping New York again.
```{r, fig.align="center", fig.width = 10, fig.height=5}
zillow_prodrop = zillow_proportion[-2]
zillow_pdmelt <- melt(zillow_prodrop, id = "Date")
names(zillow_pdmelt)[2] <- "City"
names(zillow_pdmelt)[3] <- "ProportionChange"
ggplot(zillow_pdmelt, aes(x=City, y=ProportionChange, color=City, na.rm=TRUE)) + geom_boxplot(outlier.shape=2, outlier.size=1)
```


## Anova Test and Post-hoc Tukey HSD
We will now anova test the proportional changes against each other, and see if the average change between years is different. Since we have a p-value less than our alpha, there are significant differences in changes, and thus, we look at the Post-hoc Tukey HSD. It looks like the pairs that have significantly different average proportional changes are [Detroit-LosAngeles], [Roanoke-LosAngeles], [SanFrancisco-Chicago], [Seattle-Chicago], [Richmond-Chicago], [Detroit-Houston], [Seattle-Philadelphia], [SanFrancisco-LasVegas], [Seattle-LasVegas], [Richmond-LasVegas], [Detroit-SanFrancisco], [Roanoke-SanFrancisco], [Seattle-Detroit], [Washington-Detroit], [Richmond-Detroit], [Baltimore-Seattle], [Roanoke-Seattle], and finally, [Roanoke-Richmond].
```{r, results="markup"}
pm_anova = aov(zillow_pdmelt$ProportionChange ~ zillow_pdmelt$City)
summary(pm_anova)
tukeyAoV <- TukeyHSD(pm_anova)
tukeyAoV
```

